---
title: "p8105_hw6_khs2318"
author: "Kayla Schiffer-Kane"
date: "2023-12-02"
output: 
  github_document:
    toc: true
---
```{r setup, include=FALSE}

knitr::opts_chunk$set(message = FALSE,
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

library(p8105.datasets)
library(tidyverse)
library(ggridges)
library(purrr)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

The _Washington Post_ has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository [here](https://github.com/washingtonpost/data-homicides). You can read their accompanying article [here](https://www.washingtonpost.com/graphics/2018/investigations/where-murders-go-unsolved/).

## Import and Clean Homicide Data
* Create a `city_state` variable (e.g. “Baltimore, MD”)
* Create a binary variable indicating whether the homicide is solved
* Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO -- these don't report victim race. 
* Also omit Tulsa, AL -- this is a data entry mistake. 
* Limit your analysis those for whom `victim_race` is `white` or `black`. Be sure that `victim_age` is numeric. 

```{r load_tidy}
homicide_data = read_csv('./data/homicide-data.csv') |>
  janitor::clean_names() |>
  mutate(city_state = paste(city,state, sep = ", "),
         solved = as.numeric(disposition == "Closed by arrest"),
         victim_age = as.numeric(victim_age)) |>
  filter(victim_race %in% c("White","Black")) |>
  filter(!city_state %in% 
           c("Dallas, TX","Phoenix, AZ","Kansas City, MO","Tulsa, AL")) |>
  mutate(victim_race = fct_relevel(victim_race),
         victim_sex = fct_relevel(victim_sex, "Female"))

```


## GLM Baltimore 

For the city of Baltimore, MD, use the `glm` function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of `glm` as an R object; apply the `broom::tidy` to this object; and obtain the estimate and confidence interval of the adjusted **odds ratio** for solving homicides comparing male victims to female victims keeping all other variables fixed. 

```{r baltimore_glm}
baltimore_logistic = homicide_data |> 
  filter(city_state == "Baltimore, MD") |>
  glm(solved ~ victim_age + victim_sex + victim_race, data = _, family = binomial()) 

baltimore_logistic |>
  broom::tidy(conf.int = TRUE) |>
  mutate(OR = exp(estimate)) |>
  select(term, log_OR = estimate, OR, p.value, conf.low, conf.high) |> 
  knitr::kable(digits = 3)

```

Homicides with male victims are less likely to be solved than those with female victims. 

## City-wise GLM
Now run `glm` for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. Do this within a “tidy” pipeline, making use of `purrr::map`, list columns, and `unnest` as necessary to create a dataframe with estimated ORs and CIs for each city. 
                             

```{r city_glm}
city_logistic = homicide_data |>
  nest(data = -city_state) |>
  mutate(glm = map(data, \(data) glm(solved ~ victim_sex + victim_race + victim_age,  
                                     data = data) |>
                     broom::tidy(conf.int=TRUE))) |>
  unnest(glm) |>
  mutate(OR = exp(estimate),
         conf.low = exp(conf.low),
         con.high = exp(conf.high)) |>
  select(city_state, term, OR, p.value, conf.low, conf.high) |>
  filter(term %in% c("victim_sex","victim_sexMale","victim_sexFemale"))

city_logistic |>
  knitr::kable()

```


```{r, eval = FALSE}
city_logistic |>
  group_by(city_state, term, OR) |>
  summarize(ci_lower = quantile(OR, 0.025),
         ci_upper = quantile(OR, 0.975))
```

## Plot OR and CIs 
Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

```{r plot_or_cis}
city_logistic |>
  ggplot(aes(x = reorder(city_state, OR), y = OR)) + geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

